Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-20 ms.tex
modifying figures for new compton prescription.
modifying angular momentum description.

\citet{EmsellemCappellari+:2007a} (their Fig.~2) find that $\lambda_R$
is generally $< 0.1$ on radial scales $<$ 10 per cent of the galaxy
half-light radius and that $\lambda_R$ decreases with decreasing $R$ interior to this point.  From equation (\ref{eq:rcirc}) we thus conclude that $r_{\rm
  circ} \lesssim 0.01r_{s}$, i.e.~that gas will circularize well
inside the inner sonic point of the flow ($r\lsim
0.1r_{\rm s}$; Fig.~\ref{fig:profiles}).  Angular momentum should only
become dynamically important for thermally stable flows in a region
that is causally disconnected from the outer part of the flow where
the mass accretion rate is determined.

@@

If Compton heating acts alone, i.e. $\tilde{v}_{w} = v_{w}^{C}$, then
solving equation (\ref{eq:vC}) for $\tilde{v}_{\rm w}$ yields
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 140 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 95 \eta_{0.02}^{0.34}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align} 
Compton heating can thus be significant in young stellar populations
with relatively high mass loss rates, e.g. $v_{w}^{\rm C} \gtrsim 300$
km s$^{-1}$ for $\eta \gtrsim 1$.  The accretion rate
corresponding to a state in which Compton heating self-regulates the accretion flow is given by substituting equation (\ref{eq:vC2}) into equation (\ref{eq:eddr_analytic}):
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 2\times 10^{-3} \eta_{0.02}^{0.2}T_{\rm
C,9}^{-0.8}\epsilon_{-2}^{-0.8} M_{\bullet,8}^{0.16}
&, \text{core}\\ 8\times 10^{-4} \eta_{0.02}^{0.3}T_{\rm
C,9}^{-0.7}\epsilon_{-2}^{-0.7} M_{\bullet,8}^{0.14}
&, \text{cusp}.
  \end{cases}
  \label{eq:MdotCa}
\end{align}
If the accretion luminosity $\epsilon = 10 \dot{M}/\dot{M}_{\rm edd}$ depends on the Eddington ratio as predicted by some models for
radiatively inefficient accretion flows (\citealt{Narayan&Yi95};
\citealt{Narayan+98}; \citealt{XieYuan:2012a}), then the Compton-limited accretion rate becomes
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 1.7\times 10^{-2} \eta_{0.02}^{0.12}T_{\rm
C,9}^{-0.44}M_{\bullet,8}^{0.09}
&, \text{core}\\ 8.6\times 10^{-4} \eta_{0.02}^{0.17}T_{\rm
C,9}^{-0.4} M_{\bullet,8}^{0.08}
&, \text{cusp}.
  \end{cases}
  \label{eq:MdotC}
\end{align}
We note that accretion is unlikely to be steady in cases where AGN heating dominates, due to the inherent delay between black hole feedback and the structure of the accretion flow on larger scales.   Equations (\ref{eq:MdotCa}), (\ref{eq:MdotC}) may nevertheless represent characteristic average value for the accretion rate, assuming a quasi-equilibrium state is achieved over many cycles.  
%{\bf  Total efficiency is 10 (mdot/MdotEdd). Everywhere.}

>>>>

\citet{EmsellemCappellari+:2007a} (their Fig.~2) find that $\lambda_R$
is generally $< 0.1$ on radial scales $<$ 10 per cent of the galaxy
half-light radius and that $\lambda_R$ decreases with decreasing $R$
interior to this point.  From equation (\ref{eq:rcirc}) we thus
conclude that $r_{\rm circ} \lesssim 0.01r_{s}$, i.e.~that gas will
circularize the inner sonic point of the flow ($r\gsim 0.01r_{\rm s}$;
Fig.~\ref{fig:profiles}).  Angular momentum should only become
dynamically important for thermally stable flows in a region that is
causally disconnected from the outer part of the flow. Thus, the flow
properties on large scales will generally be unaffected by angular momentum.

For $r\simeq r_{\rm circ} $ the gas will encounter a centrifugal
barrier, which means that much of the gas flowing in from larger
scales will be driven out in a conical outflow. The actual accretion
rate will be $\sim\alpha$ times the accretion rate on large scales
\citep{Li+2013}.

@@

If Compton heating acts alone, i.e. $\tilde{v}_{w} = v_{w}^{C}$, then
solving equation (\ref{eq:vC}) for $\tilde{v}_{\rm w}$ yields
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 140 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 95 \eta_{0.02}^{0.34}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align} 
Compton heating can thus be significant in young stellar populations
with relatively high mass loss rates, e.g. $v_{w}^{\rm C} \gtrsim 300$
km s$^{-1}$ for $\eta \gtrsim 1$.  The accretion rate
corresponding to a state in which Compton heating self-regulates the accretion flow is given by substituting equation (\ref{eq:vC2}) into equation (\ref{eq:eddr_analytic}):
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 2\times 10^{-3} \eta_{0.02}^{0.2}T_{\rm
C,9}^{-0.8}\epsilon_{-2}^{-0.8} M_{\bullet,8}^{0.16}
&, \text{core}\\ 8\times 10^{-4} \eta_{0.02}^{0.3}T_{\rm
C,9}^{-0.7}\epsilon_{-2}^{-0.7} M_{\bullet,8}^{0.14}
&, \text{cusp}.
  \end{cases}
  \label{eq:MdotCa}
\end{align}

\citet{Sharma+2007} calculate the radiative efficiency, $\epsilon$,
based on MHD shearing box simulations of collisionless plasmas. For
accretion rates $10^{-4} \gsim \eddr \lsim 10^{-2}$, the radiative
efficiency is $\sim 2.6\%$ {\bf AG described in more
  detail?}. Additionally, a large fraction of the $\dot{M}$ on large scales
will be driven out in outflows as the gas has some non-zero angular
momentum and a centrifugal barrier will stop it from accreting
\citet{Li+2013}. In the limit weak cooling appropriate for low
accretion rates the accretion rate on small scales will be of order
$\alpha \Mdot$. Here we adopt $\alpha=0.1$ so that our overall
efficiency is 0.26\%

\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 5.9\times 10^{-3} \eta_{0.02}^{0.12}T_{\rm
C,9}^{-0.44}M_{\bullet,8}^{0.09}
&, \text{core}\\ 2.1\times 10^{-3} \eta_{0.02}^{0.17}T_{\rm
C,9}^{-0.4} M_{\bullet,8}^{0.08}
&, \text{cusp}.
  \end{cases}
  \label{eq:MdotC}
\end{align}
We note that accretion is unlikely to be steady in cases where AGN
heating dominates, due to the inherent delay between black hole
feedback and the structure of the accretion flow on larger scales.
Equations (\ref{eq:MdotCa}), (\ref{eq:MdotC}) may nevertheless
represent characteristic average value for the accretion rate,
assuming a quasi-equilibrium state is achieved over many cycles.
%{\bf  Total efficiency is 10 (mdot/MdotEdd). Everywhere.}
